---
title: "Twitter Networks"
output: html_notebook
---

Load required libraries
```{r}
library(twitteR)
library(igraph)
library(dplyr)
```

Setup keys
```{r}
api_key <- Sys.getenv("api_key")
api_secret <- Sys.getenv("api_secret")
access_token <- Sys.getenv("access_token")
access_token_secret <- Sys.getenv("access_token_secret")
```

Setup connection
```{r}
setup_twitter_oauth(api_key, api_secret, access_token, access_token_secret)
```

Search for a keyword
```{r}
alltweets <- searchTwitter("brexitvote", n = 1500)
alltweets <- twListToDF(alltweets)
tweets <- alltweets[1:500,]
```

Split tweets into two sets: the original tweets and the retweets
We are just interested in the retweet network.
```{r}
split_point = split(tweets, tweets$isRetweet)
```

Create new variable with the retweet sender
```{r}
reTweets = mutate(split_point[['TRUE']], sender = substr(text, 5, regexpr(':', text) - 1))
```

Create edge list of the retweet network
```{r}
edge_list = as.data.frame(cbind(sender = tolower(reTweets$sender), receiver = tolower(reTweets$screenName)))
```

Add weight to the edges, based on the number of ocurrences
```{r}
edge_list = count(edge_list, sender, receiver)
```

Construct a directed graph from the edge list
```{r}
reTweets_graph <- graph_from_data_frame(d=edge_list, directed=T)
```

Plot retweet network
```{r}
par(bg="white", mar=c(1,1,1,1))
plot(reTweets_graph, 
     vertex.color="blue",
     vertex.size=(degree(reTweets_graph, mode = "in")), #sized by in-degree centrality
     vertex.label = NA,
     edge.arrow.size=0.8,
     edge.arrow.width=0.5,
     edge.width=edge_attr(reTweets_graph)$n/10, #sized by edge weight
     edge.color=hsv(h=.95, s=1, v=.7, alpha=0.5))
title("Retweet Network", cex.main=1, col.main="black")
```

By looking at the clusters, we can identify "idea"" communities.
With the in-node degree, we are also able to identify influence.

Force layouts are a class of algorithms for drawing graphs in an aesthetically-pleasing way.
They consider a force between any two nodes and determines which direction a node should move based on the sum of the force vectors.

```{r}
par(bg="white", mar=c(1,1,1,1))
plot(reTweets_graph, layout=layout.fruchterman.reingold,
     vertex.color="blue",
     vertex.size=(degree(reTweets_graph, mode = "in")), #sized by in-degree centrality
     vertex.label = NA,
     edge.arrow.size=0.8,
     edge.arrow.width=0.5,
     edge.width=edge_attr(reTweets_graph)$n/10, #sized by edge weight
     edge.color=hsv(h=.95, s=1, v=.7, alpha=0.5))
title("Retweet Network", cex.main=1, col.main="black")

```

A note about large networks:
General force algorithms can be slow for large networks.
Force Atlas 2 attempts to resolve the shortcomings of the Force algorithm by making a balance between the quality of the final layout and the speed of the computation algorithm.

https://github.com/analyxcompany/ForceAtlas2
